---
title: "Homework Assignment 1"
author: "Author: Jiyun Bak"
date: 'Assigned: Oct 24, 2020, Due Sun Nov 01, 2020 11:59PM'
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Exploration
```{r}
nyc = read.csv("nycflights.csv")

# Delete first column
nyc$X = NULL

# Factorizing variables 
nyc$origin = as.factor(nyc$origin)
nyc$tailnum = as.factor(nyc$tailnum)
nyc$month = as.factor(nyc$month)
nyc$dest = as.factor(nyc$dest)
nyc$carrier = as.factor(nyc$carrier)
nyc$flight = as.factor(nyc$flight)

```

### How many airlines are there?
```{r}
length(levels(nyc$carrier))
```

### How many flights there were by the airline with code OO?
```{r}
nyc00 = nrow(nyc[nyc$carrier == "OO",])
```

### How long is the shortest flight out of any NYC airport?
```{r}
min(nyc$air_time, na.rm=TRUE)
```

### How many flights where there by United Airlines (code: UA) on Jan 12th 2013?
```{r}
UA <- nyc[nyc$carrier == "UA" & nyc$year==2013 & nyc$month==1 & nyc$day==12, ] 
nrow(UA)

```

# Arrival Delay

### What was the average arrival delay for all airports and all airlines combined in Jan 2013?
```{r}
janflight = nyc[nyc$month == "1",]
round(mean(janflight$arr_delay, na.rm=TRUE), 2)
```

### What was the median arrival delay for all airports and all airlines combined in Jan 2013?
```{r}
round(median(janflight$arr_delay, na.rm=TRUE), 2)
```

Since the mean is greater than the median, the distribution has a longer tail in the negative direction. Thus, more flights arrive earlier than expected, however, when flights are delayed the delay is longer. 


# Airline Performance

### Calculate average arrival delays by airline
```{r}
delay = tapply(nyc$arr_delay, nyc$carrier, mean, na.rm=TRUE)
```

### Draw a Bar Plot of Average Arrival Delays for all the Airlines
```{r}
barplot(delay, main="Average Arrival Delays for all the airlines",
        xlab = "Airlines", ylab = "Avg arr delays")
```

### Which airline has the highest average arrival delay? Which airline has the smallest average arrival delay? Are there airlines that actually have negative average delay?
```{r}
# Highest average arrival delay
high <- names(delay[delay==max(delay)])

# Smallest average arrival delay
small <- names(delay[delay==min(delay)])

# Negative average delay
neg <- names(delay[delay[ ] < 0 ])

```
Highest average arrival delay airline : `r high`

Smallest average arrival delay airline : `r small`

Negative average arrival delay airline : `r neg`


# Air Gain 
```{r}
nyc$airgain <- nyc$dep_delay - nyc$arr_delay
```

### a) Explore airgain data 
```{r}
hist(nyc$airgain)

bycarrier <- tapply(nyc$airgain, nyc$carrier, mean, na.rm=TRUE)
barplot(bycarrier, main="Average airgain by the airlines",
        xlab = "Airline", ylab="Avg airgain")

byorigin <- tapply(nyc$airgain, nyc$origin, mean, na.rm=TRUE)
barplot(byorigin, main="Average airgain by Origin",
        xlab = "Origin", ylab="Avg airgain")

aggregate(airgain~origin+carrier, data=nyc, mean)
```


### b) Answer the questions - 
### i) do airlines actually gain any time when in air on average?
```{r}
carair <- aggregate(airgain~carrier, data=nyc, mean)
carair[carair$airgain < 0, ]
```
Not really. Only F9, FL, and MQ are shortening their time through flying.


### ii) Calculate average airgain for different airlines - which airlines do a better job, which do a worse job? 
```{r}
carair
#Max
carair$carrier[which.max(carair$airgain)]
#Min
carair$carrier[which.min(carair$airgain)]
```
F9 was doing a better job comparing to other airlines by shortening the average delay by 1.72. AS was doing a worse job comparing to other airlines by increasing the average delay by 15.76


# Merging Data Frames 
```{r}
library(data.table)
plane <- read.csv("planes.csv")

plane$X = NULL
plane$tailnum <- as.factor(plane$tailnum)
plane$year <- as.factor(plane$year)
plane$manufacturer <- as.factor(plane$manufacturer)
plane$model <- as.factor(plane$model)
plane$engine <- as.factor(plane$engine)
plane <- as.data.frame(plane)
```

### b) Merge the flights data and the planes data using the merge command. 

```{r}

m_data <- merge(nyc, plane, by = "tailnum")

View(m_data)
```

### c-1. Which manufacturers do make the largest&smallest average airgain plane? 
```{r}
byman <- tapply(m_data$airgain, m_data$manufacturer, mean, na.rm=TRUE)
# Highest average arrival delay
max_man <- names(byman[byman==max(byman)])
min_man <- names(byman[byman==min(byman)])
```
The largest average airgain manufacturer: `r max_man`

The smallest average airgain manufacturer: `r min_man`



### c-2. Which models do fly the average longest & shortest distance? 
```{r}
dist_model <- tapply(m_data$distance, m_data$model, mean, na.rm=TRUE)
longdist_m <- names(dist_model[dist_model==max(dist_model)])
shortdist_m <- names(dist_model[dist_model==min(dist_model)])
longdist <- dist_model[dist_model==max(dist_model)]
shortdist <- dist_model[dist_model==min(dist_model)]
longdist
shortdist
```
`r longdist_m` flies the longest distance as of `r longdist`, and `r shortdist_m` flies the shortest distance as of `r shortdist`.


### c-3. What are the top 5 most frequent destination airports flying from JFK airport?  
```{r}
m_data$cnt = 1
ori_dest <-as.data.frame(aggregate(cnt~origin+dest, data=m_data, sum))
jfk <- ori_dest[ori_dest$origin=="JFK", ]
jfk <- jfk[order(-jfk$cnt),]
jfk[1:5, 2:3]
```

### c-4. Which month does have the largest number of flights? 
```{r}
mon<-tapply(m_data$cnt, m_data$month, sum, na.rm=TRUE)
max_month <- names(mon[mon==max(mon)])
max_month
```


### c-5. Is there any relationship between the airgain and the number of engines? 
```{r}
tapply(m_data$airgain, m_data$engines, mean, na.rm=TRUE)
```
It seems that there is a linear relationship between the airgain and the number of engines until the plane has 4 engines. 
As the number of engines increases, the average airgain also increases. 